---
title: "R Code for Processing CellProfiler Data on 2D Stiffness Gradient"
subtitle: "Data for Asthma Manuscript: Dysregulated stiffness mechanosensation of airway smooth muscle cells in asthma"
author: "Yong Hwee Tan & Sebastian Amos, 2024"
output-file: "Render-for-Asthma Manuscript 2025.pdf"
execute:
  eval: false
format:
  pdf:
    toc: true
    toc-title: Contents
    number-sections: true
    colorlinks: true
    keep-tex: true
    include-in-header: 
       text: |
         \usepackage{fvextra}
         \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
         \DefineVerbatimEnvironment{OutputCode}{Verbatim}{breaklines,commandchars=\\\{\}}
editor_options: 
  chunk_output_type: inline
---

\newpage



# Introduction

This document is a render (PDF) of the R Quarto document used to process and manage all CellProfiler(TM) generated data for the Asthma manuscript; 

Asthma: Dysregulated stiffness mechanosensation of airway smooth muscle cells in asthma

# Setup, Data Directory, Preprocessing 

## Packages

Check packages and install before loading if not present.

```{r setup, include=FALSE}
install.packages("reshape2")
# Load required libraries
library(tidyverse, warn.conflicts=F, quietly=T)
library(readxl, warn.conflicts=F, quietly=T)
library(ggpubr, warn.conflicts=F, quietly=T)
library(plyr, warn.conflicts=F, quietly=T)
library(tidyr, warn.conflicts=F, quietly=T)
library(ggpmisc, warn.conflicts=F, quietly=T)
library(RColorBrewer, warn.conflicts = F, quietly = T)
library(ggsci, warn.conflicts = F, quietly = T)
library(performance, warn.conflict = F, quietly = T)
library(purrr, warn.conflict = F, quietly = T)
library(stringr, warn.conflict = F, quietly = T)
library(dplyr, warn.conflict = F, quietly = T)
library(rstatix, warn.conflict = F, quietly = T)
library(ggplot2, warn.conflict = F, quietly = T)
library(knitr, warn.conflict = F, quietly = T)
library(ggpmisc, warn.conflict = F, quietly = T)
library(forcats,  warn.conflict = F, quietly = T)

```

## Data  Directory

Define directory for all files.

```{r Directory}

outputDir <- "outputs"
dir <- getwd()
```

## Preprocessing 

Define all preprocessing functions to be repeated throughout the script

```{r}
concatenate_cellprofiler_rounds <- function(directory, pattern, file_name, grouping, outputDir){
  setwd({{directory}})
  fileList <- list.files(pattern = {{pattern}}, recursive = TRUE)
  df <- data.frame()
  for (file in fileList){
    if(nrow(df) == 0){
      df <- read_csv(file)
    }
    else {
      tempdf <- read_csv(file)
      df <- rbind(df, tempdf)
      rm(tempdf)
    }
  }
  df$index <- 1:nrow(df)
  df$type <- {{grouping}}
  df$inhibition <- ifelse(str_detect(type, "I_"), "Inhibited", "Control")
  df <- janitor::clean_names(df)
  name <- paste({{pattern}}, "Final_", sep = "_")
  filename <- paste0(name, {{grouping}},".csv")
  
  setwd(dir)
  # For Apple machines, change sep = "\\", to sep = "/" in the write_csv() function
  write_csv(df, paste({{outputDir}}, filename, sep = "/"), append = FALSE )
}
```

# Data Import

Data for asthma manuscript are defined in the following 4 groups, in the 4 chunks of data importation below. Each protein sets, A & B, are now grouped together with control and inhibited (with Y27632) data:

Set A: Channel1 - aSMA, Channel2 - YAP. Set B: Channel3 - smMHC, Channel4 - LaminA

AA - control asthma, BA - control asthma, IAA - inhibited asthma, IBA - inhibited asthma

## The Unique identifier chunk
These labels allows us to merge the dataframes from control and inhibited data. 

```{r}
addInhibition <- function(df){
  df <- df %>% 
    mutate(inhibition = ifelse(str_detect(type, "I_"), "Inhibited", "Control"))
}

```

The following 6 chunks are for Set A and Set B, of the data.

## Data Chunks for All 4 Groups
Set A Control Asthma-files coded - AA

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

aaFolder <- "data/Study2-datafiles-AA"

type <- "Asthma_A"


cytopattern <- "Cytoplasm"
name <- "Cyto_Master_"
cytodfAsthma_A <- concatenate_cellprofiler_rounds(aaFolder, cytopattern, name, type, outputDir)

cellpattern <- "Cell"
cname<- "Cell_Master_"
celldfAsthma_A <- concatenate_cellprofiler_rounds(aaFolder, cellpattern, cname, type, outputDir)

nucpattern <- "FilteredNuclei"
nname <- "Nuclei_Master_"
nucdfAsthma_A <- concatenate_cellprofiler_rounds(aaFolder, nucpattern, nname, type, outputDir)

```

Set A Control Asthma-files coded - IAA

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

aaiFolder <- "data/Study2-datafiles-InhibitedAA"
type <- "I_Asthma_A"


cytopattern <- "Cytoplasm"
name <- "Cyto_Master_"
cytodfI_Asthma_A <- concatenate_cellprofiler_rounds(aaiFolder, cytopattern, name, type, outputDir)

cellpattern <- "Cell"
cname <- "Cell_Master_"
celldfI_Asthma_A <- concatenate_cellprofiler_rounds(aaiFolder, cellpattern, cname, type, outputDir)

nucpattern <- "FilteredNuclei"
nname <- "Nuclei_Master_"
nucdfI_Asthma_A <- concatenate_cellprofiler_rounds(aaiFolder, nucpattern, nname, type, outputDir)

```

Set B Control Asthma-files coded - BA

```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}

folders <- "/Users/JoeTan/Desktop/GitHub/ASM-Analysis/data/Study2-datafiles-BA"

baFolder <- "data/Study2-datafiles-BA"

type <- "Asthma_B"


cytopattern <- "Cytoplasm"
name <- "Cyto_Master_"
cytodfAsthma_B <- concatenate_cellprofiler_rounds(baFolder, cytopattern, name, type, outputDir)

cellpattern <- "Cell"
cname <- "Cell_Master_"
celldfAsthma_B <- concatenate_cellprofiler_rounds(baFolder, cellpattern, cname, type, outputDir)

nucpattern <- "FilteredNuclei"
nname <- "Nuclei_Master_"
nucdfAsthma_B <- concatenate_cellprofiler_rounds(baFolder, nucpattern, nname, type, outputDir)

```

Set B Inhibited Asthma-files coded - IBA
 
```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results='hide'}


baiFolder <- "data/Study2-datafiles-InhibitedBA"

type <- "I_Asthma_B"


cytopattern <- "Cytoplasm"
name <- "Cyto_Master_"
cytodfI_Asthma_B <- concatenate_cellprofiler_rounds(baiFolder, cytopattern, name, type, outputDir)

cellpattern <- "Cell"
cname <- "Cell_Master_"
celldfI_Asthma_B <- concatenate_cellprofiler_rounds(baiFolder, cellpattern, cname, type, outputDir)

nucpattern <- "FilteredNuclei"
nname <- "Nuclei_Master_"
nucdfI_Asthma_B <- concatenate_cellprofiler_rounds(baiFolder, nucpattern, nname, type, outputDir)

```

# Combining Disease States

Creating a merged df which includes all disease requires an indexing system. Dfs are imported with index vals from 1:n for cell, cyto, nuc. This means there are repeats of 1:n. We needed make index values unique - chose to append disease state to index. The following groups from Set A includes; A_A, I_A_A, are to be identified and concatenated, stacked in a df. The following groups from Set B includes; B_A, I_B_A, are to be identified and concatenated, stacked in a df. 

## Concatenated Set A

Here we merge disease states (asthma, inhibited asthma) into a concatenated df for each type of df, cell, cytoplasm (cyto) or nuclei (nuc).

```{r}
# reassign index vals based on disease type

updateIndex <- function(data, string){
  data <- data %>% 
    mutate(index = paste(index, string, sep = "_"))
}

# Control ASTHMA
A_asthmaString <- "A_A"
A_asthmaList <- list(celldfAsthma_A, cytodfAsthma_A, nucdfAsthma_A)
A_asthmaIndexed <- lapply(A_asthmaList, function(df) updateIndex(df, string = A_asthmaString))

## reassign
celldfAsthma_A <- A_asthmaIndexed[[1]]
cytodfAsthma_A <- A_asthmaIndexed[[2]]
nucdfAsthma_A <- A_asthmaIndexed[[3]]

#Inhibited ASTHMA
I_A_asthmaString <- "I_A_A"
I_A_asthmaList <- list(celldfI_Asthma_A, cytodfI_Asthma_A, nucdfI_Asthma_A)
I_A_asthmaIndexed <- lapply(I_A_asthmaList, function(df) updateIndex(df, string = I_A_asthmaString))

## reassign
celldfI_Asthma_A <- I_A_asthmaIndexed[[1]]
cytodfI_Asthma_A <- I_A_asthmaIndexed[[2]]
nucdfI_Asthma_A <- I_A_asthmaIndexed[[3]]

cell_A_df <- rbind(celldfAsthma_A, celldfI_Asthma_A)

cyto_A_df <- rbind(cytodfAsthma_A, cytodfI_Asthma_A)

nuc_A_df <- rbind(nucdfAsthma_A, nucdfI_Asthma_A)

write.csv(cell_A_df, file = "outputs/cell_A_df.csv")

write.csv(cyto_A_df, file = "outputs/cyto_A_df.csv")

write.csv(nuc_A_df, file = "outputs/nuc_A_df.csv")

```

## Concatenated Set B

Here we merge disease states (asthma, inhibited asthma) into a concatenated df for each type of df, cell, cytoplasm (cyto) or nuclei (nuc).

```{r}
# reassign index vals based on disease type

updateIndex <- function(data, string){
  data <- data %>% 
    mutate(index = paste(index, string, sep = "_"))
}

# Control ASTHMA
B_asthmaString <- "B_A"
B_asthmaList <- list(celldfAsthma_B, cytodfAsthma_B, nucdfAsthma_B)
B_asthmaIndexed <- lapply(B_asthmaList, function(df) updateIndex(df, string = B_asthmaString))

## reassign
celldfAsthma_B <- B_asthmaIndexed[[1]]
cytodfAsthma_B <- B_asthmaIndexed[[2]]
nucdfAsthma_B <- B_asthmaIndexed[[3]]

# Inhibited ASTHMA
I_B_asthmaString <- "I_B_A"
I_B_asthmaList <- list(celldfI_Asthma_B, cytodfI_Asthma_B, nucdfI_Asthma_B)
I_B_asthmaIndexed <- lapply(I_B_asthmaList, function(df) updateIndex(df, string = I_B_asthmaString))

## reassign
celldfI_Asthma_B <- I_B_asthmaIndexed[[1]]
cytodfI_Asthma_B <- I_B_asthmaIndexed[[2]]
nucdfI_Asthma_B <- I_B_asthmaIndexed[[3]]

cell_B_df <- rbind(celldfAsthma_B, celldfI_Asthma_B)

cyto_B_df <- rbind( cytodfAsthma_B, cytodfI_Asthma_B)

nuc_B_df <- rbind( nucdfAsthma_B, nucdfI_Asthma_B)

write.csv(cell_B_df, file = "outputs/cell_B_df.csv")

write.csv(cyto_B_df, file = "outputs/cyto_B_df.csv")

write.csv(nuc_B_df, file = "outputs/nuc_B_df.csv")

```

# Modifying Each Data Frame Before Combination 

In the following section, we create new columns and modify column names for both Set A and B, 

```{r echo=TRUE, warning=FALSE, paged.print=FALSE, results='hide'}

# CYTOPLASM
cyto_A_df <- cyto_A_df %>% 
  mutate(aSMA_cyto = intensity_integrated_intensity_channel1) %>%
  mutate(YAP_cyto = intensity_integrated_intensity_channel2) %>%
  mutate(aSMA_cyto_rel = aSMA_cyto / area_shape_area, .after=aSMA_cyto) %>%
  mutate(YAP_cyto_rel = YAP_cyto / area_shape_area, .after=YAP_cyto) 

#Set B holds data for smMHC-Channel3, LaminA-Channel4.*only channel3 requires changes in cyto_Bdf.
cyto_B_df <- cyto_B_df %>%
  mutate(smMHC_cyto = intensity_integrated_intensity_channel3) %>%
  mutate(smMHC_cyto_rel = smMHC_cyto / area_shape_area, .after=smMHC_cyto)
  
#CELLS
cell_A_df <- cell_A_df %>% # update the celldf with the following:
  mutate(cell_aspect_ratio = area_shape_major_axis_length / area_shape_minor_axis_length, .after=area_shape_major_axis_length) %>% 
  mutate(aSMA_cell = intensity_integrated_intensity_channel1) %>% 
  mutate(aSMA_cell_rel = aSMA_cell / area_shape_area, .after=aSMA_cell) %>% 
  mutate(cell_form_factor = area_shape_form_factor) %>% 
  mutate(cell_orientation = area_shape_orientation) 

cell_B_df <- cell_B_df %>% # update the celldf with the following. 
  mutate(cell_aspect_ratio = area_shape_major_axis_length / area_shape_minor_axis_length, .after=area_shape_major_axis_length) %>% 
  mutate(smMHC_cell = intensity_integrated_intensity_channel3) %>% 
  mutate(smMHC_cell_rel = smMHC_cell / area_shape_area, .after=smMHC_cell) %>%  
  mutate(cell_form_factor = area_shape_form_factor) %>% 
  mutate(cell_orientation = area_shape_orientation) 

#NUCLEI
nuc_A_df <- nuc_A_df %>% 
  mutate(YAP_nuc = intensity_integrated_intensity_channel2) %>%
  mutate(YAP_nuc_rel = YAP_nuc / area_shape_area, .after=YAP_nuc) %>%
  mutate(nuc_aspect_ratio =  area_shape_major_axis_length / area_shape_minor_axis_length, .after=area_shape_major_axis_length) %>%
  mutate(aSMA_nuc = intensity_integrated_intensity_channel1) %>%
  mutate(aSMA_nuc_rel = aSMA_nuc / area_shape_area, .after=aSMA_nuc) %>%
  mutate(nuc_form_factor = area_shape_form_factor) %>%
  mutate(nuc_aspect_ratio = area_shape_major_axis_length / area_shape_minor_axis_length, .after=area_shape_major_axis_length) %>% 
  mutate(nuc_orientation = area_shape_orientation) 

nuc_B_df <- nuc_B_df %>% 
  mutate(LaminA = intensity_integrated_intensity_channel4) %>%
  mutate(LaminA_rel = LaminA / area_shape_area, .after=LaminA) %>%
  mutate(smMHC_nuc =  intensity_integrated_intensity_channel3) %>%
  mutate(smMHC_nuc_rel = smMHC_nuc / area_shape_area, .after=smMHC_nuc) %>%
  mutate(nuc_form_factor = area_shape_form_factor) %>%
  mutate(nuc_aspect_ratio = area_shape_major_axis_length / area_shape_minor_axis_length, .after=area_shape_major_axis_length) %>% 
  mutate(nuc_orientation = area_shape_orientation) 

```

## Merge Data Frames

We create a unique index column in the concatenation function so that we have a common ID value between data frames we want to join together. First, the correct data (with every column having a unique name) will be selected before merging - note: the common variable between data frames, index, is selected in each case. ImageNumber is also selected in the cell df so it can be decoded into stiffness next. In this chunk, the mutate() function is used in place of the rename() function - the outcome is the same.

```{r}
# CYTO 
cyto_A <- cyto_A_df %>%
  mutate(cyto_area = area_shape_area, cyto_number = object_number) %>%
  select(cyto_area, cyto_number, aSMA_cyto, aSMA_cyto_rel, YAP_cyto, YAP_cyto_rel, index, type, inhibition)

cyto_B <- cyto_B_df %>%
  mutate(cyto_area = area_shape_area, cyto_number = object_number) %>%
  select(cyto_area, cyto_number, smMHC_cyto, smMHC_cyto_rel, index, type, inhibition)

# CELLS
cell_A <- cell_A_df %>% 
  mutate(cell_area = area_shape_area, cell_number = object_number) %>%
  select(cell_area , cell_number, cell_aspect_ratio, cell_form_factor, cell_orientation, aSMA_cell, aSMA_cell_rel, index, image_number, type, inhibition) 

cell_B <- cell_B_df %>%
  mutate(cell_area = area_shape_area, cell_number = object_number) %>%
  select(cell_area , cell_number, cell_aspect_ratio, cell_form_factor, cell_orientation, smMHC_cell, smMHC_cell_rel, index, image_number, type, inhibition) 
  
# NUCLEI
nuc_A <- nuc_A_df %>% 
  mutate(nuc_area = area_shape_area, nuc_number = object_number) %>% 
  select(nuc_area, nuc_number, nuc_aspect_ratio, nuc_form_factor, nuc_orientation, aSMA_nuc, aSMA_nuc_rel, YAP_nuc, YAP_nuc_rel, index, type, inhibition) 

nuc_B <- nuc_B_df %>%
  mutate(nuc_area = area_shape_area, nuc_number = object_number) %>% 
  select(nuc_area, nuc_number, nuc_aspect_ratio, nuc_form_factor, nuc_orientation,smMHC_nuc, smMHC_nuc_rel, LaminA, LaminA_rel, index, type, inhibition) 
 

# Bind the various data frames together, combinig only the relevant information.
#For Set A
dfDistinctCellA <- cell_A %>% 
  left_join(nuc_A, by = "index" , "type")

dfA <- dfDistinctCellA %>% 
  left_join(cyto_A, by = "index", "type")

#For Set B
dfDistinctCellB <- cell_B %>% 
  left_join(nuc_B, by = "index", "type")

dfB <- dfDistinctCellB %>% 
  left_join(cyto_B, by = "index", "type")

```

# Modifying the Final Data Frame-Protein Markers
Now we have a combined data frame for Set A "dfA", and Set B "dfB" that includes Control and Inhibition data.

Before we do any analysis, we decode our stiffness values (which are currently 1:8) into the actual stiffness measurements. This will allow us to factorise or group by stiffness and plot with accurate values instead of placeholders. Then, we filter the data based on some conditions (i.e. object area).

The following categorising chunks for Set A and B provide the df's for the four protein markers in our studies; aSMA, YAP, Lamin-A and sm-MHC. 

## Categorising Set A according to stiffness 
First, we modify the df by stiffness. Each stiffness range was imaged on four locations in a row. We then modify and rename each of the columns in anticipation for the categorizing we will apply in Graphpad Prism.

```{r}
# MODIFY THE FINAL SET A DATA FRAME - REDONE for BOTH CONTROLS AND INHIBITION
dfA <- dfA %>% 
   mutate(stiffness = case_when(image_number %in% c( 1 , 2 , 3 , 4 ) ~ 2.6,
                               image_number %in% c( 5 , 6 , 7 , 8) ~ 5,
                               image_number %in% c( 9 , 10 , 11 , 12 )  ~ 9,
                               image_number %in% c( 13 , 14, 15 , 16 ) ~ 13.5,
                               image_number %in% c( 17 , 18 , 19 , 20 )  ~ 18.5,
                               image_number %in% c( 21 , 22 , 23 , 24 )  ~ 23,
                               image_number %in% c( 25 , 26 , 27 , 28 ) ~ 27,
                               image_number %in% c( 29 , 30 , 31 , 32 )  ~ 31))

dfFacA <- dfA %>% 
  mutate(stiffness = as.factor(stiffness), type = as.factor(type)) %>% 
  mutate(nuc_cellsize_ratio = nuc_area/cell_area, .before=cell_area) %>%
  mutate(YAP_nuccyto = YAP_nuc/YAP_cyto, .after=YAP_nuc) %>% 
  mutate(YAP_nuccyto_rel = YAP_nuc_rel/YAP_cyto_rel, .after=YAP_nuc_rel) %>%
  mutate(aSMA_nuccyto = aSMA_nuc/aSMA_cyto, .after=aSMA_nuc) %>%
  mutate(aSMA_nuccyto_rel = aSMA_nuc_rel/aSMA_cyto_rel, .after=aSMA_nuc_rel) %>%
  mutate(cell_areamicrons = cell_area*0.436464752, .after=cell_area) %>% # Application of conversion factor, according to single pixel area in microns squared (cellprofiler provides area quantified by total pixel number)
  mutate(nuc_areamicrons = nuc_area*0.436464752, .after=nuc_area) %>%
  mutate(cyto_areamicrons = cyto_area*0.436464752, .after=cyto_area) %>%
  filter(nuc_cellsize_ratio != 1)  %>% # Filters out rows where nuc_cellsize_ratio equals 1
  mutate(type = recode(type,
                             "Asthma_A" = "AsthmaA"))
  
# Filter based on the type values
dfAsthmaA <- dfFacA %>% filter(str_detect(type, "AsthmaA")) 
dfInhibAsthmaA <- dfFacA %>% filter(str_detect(type, "I_Asthma_A")) 

# Save each filtered data frame as a separate CSV file
write.csv(dfAsthmaA, file = "outputs/dfAsthmaA.csv", row.names = FALSE)
write.csv(dfInhibAsthmaA, file = "outputs/dfInhibAsthmaA.csv", row.names = FALSE)

# Write a CSV for the full data frame.
write.csv(dfFacA, file = "outputs/dfFacA.csv")

```

## Categorising Set B according to stiffness 

First, we modify the df by stiffiness. Each stiffness range was imaged on four locations in a row. We then modify and rename each of the columns in anticipation for the categorizing we will apply in Graphpad Prism.

```{r}
# MODIFY THE FINAL SET B DATA FRAME - REDONE for BOTH CONTROLS AND INHIBITION
dfB <- dfB %>% 
   mutate(stiffness = case_when(image_number %in% c( 1 , 2 , 3 , 4 ) ~ 2.6,
                               image_number %in% c( 5 , 6 , 7 , 8) ~ 5,
                               image_number %in% c( 9 , 10 , 11 , 12 )  ~ 9,
                               image_number %in% c( 13 , 14, 15 , 16 ) ~ 13.5,
                               image_number %in% c( 17 , 18 , 19 , 20 )  ~ 18.5,
                               image_number %in% c( 21 , 22 , 23 , 24 )  ~ 23,
                               image_number %in% c( 25 , 26 , 27 , 28 ) ~ 27,
                               image_number %in% c( 29 , 30 , 31 , 32 )  ~ 31))

dfFacB <- dfB %>% 
  mutate(stiffness = as.factor(stiffness), type = as.factor(type)) %>% 
  mutate(nuc_cellsize_ratio = nuc_area/cell_area, .before=cell_area) %>%
  mutate(LaminA_rel_cell = LaminA/cell_area, .after=LaminA) %>%
  mutate(smMHC_rel_cell = smMHC_cell/cell_area, .after=smMHC_cell) %>%
  mutate(cell_areamicrons = cell_area*0.436464752, .after=cell_area) %>% # Application of conversion factor, according to single pixel area in microns squared (cellprofiler provides area quantified by total pixel number)
  mutate(nuc_areamicrons = nuc_area*0.436464752, .after=nuc_area) %>%
  mutate(cyto_areamicrons = cyto_area*0.436464752, .after=cyto_area) %>%
  filter(nuc_cellsize_ratio != 1)  %>% # Filters out rows where nuc_cellsize_ratio equals 1
  filter(smMHC_rel_cell != 1)  %>% # Filters out rows where nuc_cellsize_ratio equals 1
  mutate(type = recode(type,
                             "Asthma_B" = "AsthmaB"))
# Now we filter based on the type values
dfAsthmaB <- dfFacB %>% filter(str_detect(type, "AsthmaB"))
dfInhibAsthmaB <- dfFacB %>% filter(str_detect(type, "I_Asthma_B"))

# Save each filtered data frame as a separate CSV file
write.csv(dfAsthmaB, file = "outputs/dfAsthmaB.csv", row.names = FALSE)
write.csv(dfInhibAsthmaB, file = "outputs/dfInhibAsthmaB.csv", row.names = FALSE)

# Write a CSV for the full data frame.
write.csv(dfFacB, file = "outputs/dfFacB.csv")

```

# Modify Final Data Frame-Morphology Assessment
In this section, we combine morphological data from both Set A and B.

```{r}
# Collect morphology columns into new dfs
dfAMorpho <- dfA %>% 
  select(index, image_number, cell_area, cell_number, cell_aspect_ratio, cell_form_factor, cell_orientation, nuc_area, nuc_number, nuc_aspect_ratio, nuc_form_factor,nuc_orientation,cyto_area, cyto_number, type, inhibition, stiffness)

dfBMorpho <- dfB %>% 
  select(index, image_number, cell_area, cell_number, cell_aspect_ratio, cell_form_factor, cell_orientation, nuc_area, nuc_number, nuc_aspect_ratio, nuc_form_factor, nuc_orientation, cyto_area, cyto_number, type, inhibition, stiffness)

# Bind them together
dfMorpho <- rbind(dfAMorpho, dfBMorpho)

# Creating a new column, to allocate disease type as a merged group including both set A and B data

write.csv(dfMorpho, file = "outputs/dfMorpho.csv")

```

## Combine Set A and B morphology 
Here, we create a new column to allow combination and indexing of both Sets.

```{r}
# Creating a new colum in dfMorpho
dfMorphomerge <- dfMorpho %>% 
mutate(mergedtype = str_replace(type, "_[A-B]$", "_"))

write.csv(dfMorphomerge, file = "outputs/dfMorphomerge.csv")

```

## Categorizing Data to Stiffness 
We apply the segregation of dfMorphomerged, into dfFacMorphomerge.

```{r}
dfMorphomerge <- dfMorphomerge %>% 
   mutate(stiffness = case_when(image_number %in% c( 1 , 2 , 3 , 4 ) ~ 2.6,
                               image_number %in% c( 5 , 6 , 7 , 8) ~ 5,
                               image_number %in% c( 9 , 10 , 11 , 12 )  ~ 9,
                               image_number %in% c( 13 , 14, 15 , 16 ) ~ 13.5,
                               image_number %in% c( 17 , 18 , 19 , 20 )  ~ 18.5,
                               image_number %in% c( 21 , 22 , 23 , 24 )  ~ 23,
                               image_number %in% c( 25 , 26 , 27 , 28 ) ~ 27,
                               image_number %in% c( 29 , 30 , 31 , 32 )  ~ 31))

  dfFacMorphomerge <- dfMorphomerge %>% 
  mutate(stiffness = as.factor(stiffness), type = as.factor(type)) %>% 
  mutate(nuc_cellsize_ratio = nuc_area/cell_area, .before=cell_area) %>%
  mutate(nuclearrelcell_area = nuc_area/cell_area, .after=cell_area) %>% 
  mutate(cell_areamicrons = cell_area*0.436464752, .after=cell_area) %>% # Application of conversion factor, according to single pixel area in microns squared (cellprofiler provides area quantified by total pixel number)
  mutate(nuc_areamicrons = nuc_area*0.436464752, .after=nuc_area) %>%
  mutate(cyto_areamicrons = cyto_area*0.436464752, .after=cyto_area) %>%
  filter(nuc_cellsize_ratio != 1)  %>% # Filters out rows where nuc_cellsize_ratio equals 1
  mutate(mergedtype = recode(mergedtype,
                             "Asthma_" = "Asthma", "I_Asthma_" = "Inhibited Asthma"))

# Write the entire dataset to a single CSV file
write.csv(dfFacMorphomerge, file = "outputs/dfFacMorphomerge.csv", row.names = FALSE)

df_AsthmaMerge <- dfFacMorphomerge %>% filter(str_detect(mergedtype, "Asthma"))
df_InhibitedAsthmaMerge <- dfFacMorphomerge %>% filter(str_detect(mergedtype, "Inhibited Asthma"))

# Split the data based on 'mergedtype' and save each subset into a separate CSV
# Filter for Asthma
df_AsthmaMerge <- dfFacMorphomerge %>% filter(mergedtype == "Asthma")
write.csv(df_AsthmaMerge, file = "outputs/dfFacMorphomerge_Asthma.csv", row.names = FALSE)

df_InhibitedAsthmaMerge <- dfFacMorphomerge %>% filter(mergedtype == "Inhibited Asthma")
write.csv(df_InhibitedAsthmaMerge, file = "outputs/dfFacMorphomerge_InhibitedAsthma.csv", row.names = FALSE)

```

This concludes the markdown. All data processed can now be plotted using GraphPad Prism. 
